Title?

### 2016-07-29 draft.tex
tiny change to footnote
Went through II.B - II.D.

At this point the calculation diverges from the axion bremsstrahlung calculation, since in the kinematic region of interest for supernova explosions, the mass of the $A'$ boson is not necessarily negligible. The kinematic relations are thus \bea 
p_1 \cdot p_2 &=& M_N^2 - \frac{l^2}{2} - \frac{k^2}{2} + p_2 \cdot q_a,\\
p_1 \cdot p_3 &=& M_N^2 + k \cdot l - \frac{k^2}{2} + p_3 \cdot q_a,\\  
p_1 \cdot p_4 &=& k \cdot l + M_N^2 - \frac{l^2}{2} + p_4 \cdot q_a, \\
p_2 \cdot p_3 &=& M_N^2 - \frac{l^2}{2}, \\ 
p_2 \cdot p_4 &=& M_N^2 - \frac{k^2}{2},\\
p_3 \cdot p_4 &=& k \cdot l + M_N^2 - \frac{l^2 + k^2}{2}.
\eea
Unfortunately, with these relations it does not seem to be possible to simplify the amplitude into an analytically usable form without making unjustified approximations, and so the final expression for the spin-averaged squared matrix element, $ \mathcal{M}^2_{pp}$, contains some two hundred and fifty terms, and thus we do not reproduce that result here.\footnote{At this point we diverge from Ref.~\cite{Dent:2012mx} for two reasons.  One, they use the kinematics as if the $A'$ were massless, as in the axion calculation, and independent of that choice, their squared matrix element does not respect the correct exchange symmetry, as can be seen by inspection of their Eq. (A.3), with the claim that only $C_k$ appearing in that equation is nonzero.}

@@

\subsection{Streaming limit}
The first and simplest bound that may obtained arises from assuming that all produced particles leave the supernova, carrying their energy with them. The constraint is derived simply by requiring that the energy loss through this cooling channel be roughly less than the cooling from neutrino emission; any greater, and it would have an observable effect on supernova cooling.

The actual quantity of interest therefore is the rate of energy emission through dark bosons. From the spin-summed amplitudes already found this is obtained by integrating over the phase space, and adding a factor of energy

\beq
Q_i = \int (2\pi)^4 E_A \sum_{spins} \mathcal{M}^2_i f(p_1) f(p_2)\delta(p_1+p_2-p_3-p_4-q_a) dLIPS.
\eeq
Here $ E_A $ is the energy of the emitted boson, and $ f(p) $ are the initial occupation numbers. The nucleons in the core are comfortably non-degenerate and non-relativistic, so  the Pauli blocking factor is omitted and the Maxwell-Boltzmann distribution used  
\beq
	f(p) =  \frac{n_b}{2} (\frac{2 \pi}{M_N T})^{3/2} e^{-\frac{\textbf{p}^2} {2 M_N T}}.
\eeq

This integration is performed numerically to obtain $ Q_i $, which is the rate of energy emission per unit volume associated with either the pn or pp process. To obtain the dark gauge boson luminosity, it is assumed that production takes place in a core volume of radius 1 km, so $ \mathcal{L}_A = V(Q_{pp} + Q_{pn}) $. $ \mathcal{L}_A $ is a complicated function of $ m_A $ and $ T $, but is simply proportional to the square of the parameter $ \epsilon $; pulling that out of the expression gives a constraint 
\beq
\epsilon^2 I_A(m_A, T) \le \mathcal{L}_\nu
\eeq
or a bound on the coupling of 
\beq 
\epsilon \le \sqrt{\frac{4.1 * 10^{37} MeV^2}{I_A(m_A, T)}} 
\eeq 
and the exclusion region is generated by varying the mass.

\subsection{Decay limit}
The constraint from the streaming limit above provides an upper bound on the allowed coupling and lower bound on the excluded coupling by considering purely production. Naively it might be expected that this suffices, that all higher couplings are excluded. However, in order to function as a cooling channel, obviously enough of the produced light gauge bosons must escape the supernova. As the coupling increases, so do processes that prevent this,and so the excluded region has an upper bound, above which the luminosity again drops below the neutrino luminosity. The first such limit may be found by considering decay of the dark bosons into Standard Model particles, and assuming that these SM particles are trapped in the supernova core and contribute nothing to the cooling. The dark boson has a typical lifetime of 
\beq
l = \frac{3 E_{A}}{N_{eff} m_A^2 \epsilon^2},
\eeq
and so the fraction escaping the supernova is given by 
\beq
e^{\frac{r_{decay}}{l}} = e^{\frac{r_{decay} N_{eff} m_A^2 \epsilon^2}{3 E_A}}.
\eeq
To take this into account, the above exponential factor is simply appended to the phase space integrand, and the calculation then proceeds as before. The limit is derived from the same equation, with the only complication being the fact that $ I_A $ is now a function of $ \epsilon $, in addition to $ m_A, T $. This makes the numerical calculation slightly more complicated, but otherwise changes nothing of importance. 
\beq
\epsilon \le \sqrt{\frac{4.1 * 10^{37} MeV^2}{I_A(m_A, T, \epsilon)}}.
\eeq

\subsection{Trapping limit}
The second constraint that produces an upper bound on the excluded region comes from considering trapping of dark bosons within the supernova. With a large enough coupling, the new particles will thermalize and then be emitted from a spherical shell. In this case the luminosity is given simply by the Steffan-Boltzmann law
\beq
\mathcal{L}_t  = 4\pi r^2 T_A^4 \sigma,
\eeq
where $ r $ is now the radius of the emitting shell and $ T_A $ its temperature. These must typically be calculated, but for $ r $ it is safe to assume a value of 10 km, since the density of the supernova drops drastically around that point. After taking that value for r, the bound on the luminosity translates into a bound on $ T_A $ 
\beq
T_A \le 9.586 MeV.
\eeq
That bound can then be translated into the desired bound on the coupling as a function of mass by assuming that the particles are emitted from an optical depth $ \tau = \frac{2}{3} $, and finding the temperature that corresponds to that optical depth. This is a somewhat involved calculation.
First, one needs a model for the density and temperature in the supernova. Following \cite{dent}, we assume 
\bea
\rho &=& \rho_p (\frac{R}{r})^2,\\
T &=& T_R (\frac{R}{r})^{n/3},
\eea
with $  \rho_p  = 3*10^{14} g/cm^3, T_R = 30$ MeV, and taking $ n = 5 $. The optical depth is given by 
\beq
\tau = \int_{r_x}^{\infty} \kappa \rho dr,
\eeq
where $ \kappa$ is the opacity.  To find the opacity, we start from the reduced mean Rosseland opacity 
\beq
\frac{1}{\kappa \rho} = \int_{m_x}^{\infty} \frac{15}{4 \pi^4 T^5} \frac{E_A^2 e^{E_A/T} \sqrt{E_A^2 - m_A^2}}{(e^{E_A/T}-1)^2} l_A dE_A,
\eeq
where $ l_A $ is the mean free path.

The inverse mean free path can readily be obtained by modifying $ Q_i $, the expression for the energy loss rate, as follows: removing the factor of $ E_A $ and the phase space integral over $ q_A $, and adding a factor of $ e^{E_A/T} $ for detailed balance. This gives the inverse mean free path as a function of mass and coupling. Again the required integration is performed numerically. This then allows the calculation of $ \kappa_x $. The inverse opacities for the pn and pp processes add, giving the total opacity $ \kappa^{-1} = \kappa_{pp}^{-1} + \kappa_{pn}^{-1} $

>>>>

At this point the calculation diverges from the axion bremsstrahlung calculation, since in the kinematic region of interest for supernova explosions, the mass of the $A'$ boson is not necessarily negligible. The kinematic relations are thus \bea 
p_1 \cdot p_2 &=& M_N^2 - \frac{l^2}{2} - \frac{k^2}{2} + p_2 \cdot q_a,\\
p_1 \cdot p_3 &=& M_N^2 + k \cdot l - \frac{k^2}{2} + p_3 \cdot q_a,\\  
p_1 \cdot p_4 &=& k \cdot l + M_N^2 - \frac{l^2}{2} + p_4 \cdot q_a, \\
p_2 \cdot p_3 &=& M_N^2 - \frac{l^2}{2}, \\ 
p_2 \cdot p_4 &=& M_N^2 - \frac{k^2}{2},\\
p_3 \cdot p_4 &=& k \cdot l + M_N^2 - \frac{l^2 + k^2}{2}.
\eea
Unfortunately, with these relations it does not seem to be possible to simplify the amplitude into an analytically usable form without making unjustified approximations, and so the final expression for the spin-averaged squared matrix element, $ \mathcal{M}^2_{pp}$, contains some two hundred and fifty terms, and thus we do not reproduce that result here.\footnote{At this point we diverge from Ref.~\cite{Dent:2012mx} for two reasons.  One, they use the kinematics as if the $A'$ were massless, as in the axion calculation, and independent of that choice, their squared matrix element does not respect the correct exchange symmetry, as can be seen by inspection of their Eq. (A.3), with their claim that only the coefficient $C_k$ appearing in that equation is nonzero.}

@@

\subsection{Streaming limit}
The first and simplest bound that may be obtained arises from assuming that all produced particles leave the supernova, carrying their energy with them. The constraint is derived simply by requiring that the energy loss through this cooling channel be roughly less than the cooling from neutrino emission; any greater, and it would have an observable effect on supernova cooling.

The actual quantity of interest therefore is the rate of energy emission through dark bosons. From the spin-summed squared amplitudes already given, this is obtained by integrating over the phase space, and adding a factor of energy
\beq
Q_i = \int (2\pi)^4 E_A \sum_{spins} \mathcal{M}^2_i f(p_1) f(p_2)\delta(p_1+p_2-p_3-p_4-q_a) d\Pi,
\eeq
where $d\Pi$ is the Lorentz-invariant phase space, $ E_A $ is the energy of the emitted boson, and $ f(p) $ are the initial occupation numbers. The nucleons in the core are comfortably non-degenerate and non-relativistic, so  the Pauli blocking factor is omitted and the Maxwell-Boltzmann distribution is used  
\beq
f(p) =  \frac{n_b}{2} (\frac{2 \pi}{M_N T})^{3/2} e^{-\frac{\textbf{p}^2} {2 M_N T}}.
\eeq

The integration is performed numerically to obtain $ Q_i$, which is the rate of energy emission per unit volume associated with either the $pn$ or $pp$ process. To obtain the dark gauge boson luminosity, it is assumed that production takes place in a core volume $V$ of radius 1 km, giving $ \mathcal{L}_A = V(Q_{pp} + Q_{pn}) $. The luminosity $\mathcal{L}_A $ is a complicated function of $ m_A $ and $ T $, but is simply proportional to the square of the parameter $ \epsilon $.  Pulling that out of the expression gives a constraint 
\beq
\epsilon^2 I_A(m_A, T) \le \mathcal{L}_\nu
\eeq
or a bound on the coupling of 
\beq 
\epsilon \le \sqrt{\frac{4.1 \times 10^{37} MeV^2}{I_A(m_A, T)}} 
\eeq 
and the exclusion region is generated by varying the mass.

\subsection{Decay limit}
The constraint from the streaming limit above provides an upper bound on the allowed coupling (or lower bound on the excluded coupling) by considering the production on the $A'$. Naively it might be expected that this suffices, in that all higher couplings are excluded. However, in order to function as a cooling channel, enough of the produced dark gauge bosons must escape the supernova. As the coupling increases, so do processes that prevent the escape, and so the excluded region has an upper bound, above which the luminosity again drops below the neutrino luminosity. The first such limit may be found by considering decay of the dark bosons into Standard Model particles, and assuming that these SM particles are trapped in the supernova core and contribute nothing to the cooling. The dark boson has a typical lifetime of 
\beq
l = \frac{3 E_{A}}{N_{eff} m_A^2 \epsilon^2},
\eeq
and so the fraction escaping the supernova before decaying is given by 
\beq
e^{\frac{r_{decay}}{l}} = e^{\frac{r_{decay} N_{eff} m_A^2 \epsilon^2}{3 E_A}}.
\eeq
To take this into account, the above exponential factor is simply appended to the phase space integrand, and the calculation then proceeds as before. The limit is derived from the same equation, with the only complication being the fact that $ I_A $ is now a function of $ \epsilon $, in addition to $ m_A, T $. This makes the numerical calculation slightly more complicated, but otherwise changes nothing of importance, with now the constraint a transcendental equation,  
\beq
\epsilon \le \sqrt{\frac{4.1 * 10^{37} MeV^2}{I_A(m_A, T, \epsilon)}}.
\eeq

\subsection{Trapping limit}
The second constraint that produces an upper bound on the excluded region comes from considering trapping of dark bosons within the supernova. With a large enough coupling, the new particles will thermalize and then they will be emitted from a spherical shell.  In this case the luminosity is given simply by the Steffan-Boltzmann law
\beq
\mathcal{L}_t  = 4\pi r^2 T_A^4 \sigma,
\eeq
where $ r $ is now the radius of the emitting shell and $ T_A $ its temperature. We can estimate $ r = 10$ km, since the density of the supernova drops drastically around that point. After taking that value for r, the bound on the luminosity translates into a bound on $ T_A $ 
\beq
T_A \le 9.586\rm\ MeV.
\eeq
That bound can then be translated into the desired bound on the coupling as a function of mass by assuming that the particles are emitted from an optical depth $ \tau = 2/3 $, and finding the temperature that corresponds to that optical depth. This is a somewhat involved calculation.
First, one needs a model for the density and temperature in the supernova. Following \cite{Dent:2012mx}, we assume 
\bea
\rho &=& \rho_p \left(\frac{R}{r}\right)^2,\\
T &=& T_R \left(\frac{R}{r}\right)^{n/3},
\eea
with $  \rho_p  = 3\times10^{14}\ {\rm g/cm}^3, T_R = 30$ MeV, and taking $ n = 5 $. The optical depth is given by 
\beq
\tau = \int_{r_x}^{\infty} \kappa \rho dr,
\eeq
where $ \kappa$ is the opacity.  To find the opacity, we start from the reduced mean Rosseland opacity 
\beq
\frac{1}{\kappa \rho} = \int_{m_x}^{\infty} \frac{15}{4 \pi^4 T^5} \frac{E_A^2 e^{E_A/T} \sqrt{E_A^2 - m_A^2}}{(e^{E_A/T}-1)^2} l_A dE_A,
\eeq
where $ l_A $ is the mean free path.

The inverse mean free path can readily be obtained by modifying $ Q_i $, the expression for the energy loss rate, as follows: removing the factor of $ E_A $ and the phase space integral over $ q_A $, and adding a factor of $ e^{E_A/T} $ for detailed balance. This gives the inverse mean free path as a function of mass and coupling. Again the required integration is performed numerically. This then allows the calculation of $ \kappa_x $. The inverse opacities for the $pn$ and $pp$ processes add, giving the total opacity $ \kappa^{-1} = \kappa_{pp}^{-1} + \kappa_{pn}^{-1} $
